### 2013-01-16 paper.tex
submmited version

% \documentclass[12pt]{article}
\documentclass[a4paper,11pt,reqno]{amsart}

% This first part of the file is called the PREAMBLE. It includes
% customizations and command definitions. The preamble is everything
% between \documentclass and \begin{document}.

\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
\usepackage{graphicx}              % to include figures
\usepackage{amsmath}               % great math stuff
\usepackage{amsfonts}              % for blackboard bold, etc
\usepackage{amsthm}                % better theorem environments
\usepackage{amssymb}
\usepackage[all]{xy}

% various theorems, numbered by section

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{defn}[thm]{Definition}
\newtheorem{asser}[thm]{Assertion}
\newtheorem{rmk}[thm]{Remark}

\DeclareMathOperator{\id}{id}

\newcommand{\bd}[1]{\mathbf{#1}}  % for bolding symbols
\newcommand{\RR}{\mathbb{R}}      % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}}      % for Integers
\newcommand{\CC}{\mathbb{C}}
\newcommand{\col}[1]{\left[\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}
\newcommand{\vect}[1]{\vec{#1}}

\begin{document}

\nocite{*}

\title{Algebraic Topology Final Report: Morse Theory}

\author{Shih-Kai Chiu \\
r01221031}

\maketitle

\begin{abstract}
  This report covers roughly the part I of J. Milnor's book\cite{Mil}. The main
  result is the theorem: every smooth manifoldhas the same homotopy type of a
  CW-complex. We prove this theorem by developing the concept of non-degenerate
  smooth functions, their index at critical points and the correspondence
  between the index and the dimension of attaching cells. After the main
  theorem, the existence of non-degenerate smooth functions on manifolds is
  established, and as an application, a proof of the Lefschetz hyperplane
  theorem is given.
\end{abstract}

\section{Basic Definitions and Lemmas}

\subsection{Attaching Cells}

Let $Y$ be any topological space, and let
\begin{equation}
  e^k = \{ x \in \RR^k | \left| x \right| < 1 \}
\end{equation}
be the k-cell. The boundary
\begin{equation}
  \dot{e}^k = \{ x \in \RR^k | \left| x \right| = 1 \}
\end{equation}
is the (k-1)-sphere, $S^{k-1}$. Specifically, we put $e^0$ as a point and set
$\dot{e}^0 = \emptyset$. We call a continuous map $g: S^{k-1} \to Y$ an
attaching map and obtain a new topological space
\begin{equation}
  Y \cup_g e^k
\end{equation}
by identifying each $x \in \dot{e}^k$ with $g(x) \in Y$.

\subsection{Non-degenerate critical points}

Let $M^n$ be a smooth manifold of dimension $n$, and let $f: M \to \RR$ be a
smooth function on $M$. We recall that a point $x \in M$ is called a critical
point of $f$ if the differential map $ df_p : T_pM \to \RR $ is singular. This
is equivalent to $\frac{\partial f}{\partial x^i} (p) = 0$ for all $i = 1,
.. ,k$ and any local coordinate system $(x^1, .., x^k)$, since the map $df_p$ in
local coordinates is a row matrix with a non-empty kernel.

\begin{defn}
  Let $p \in M$ and let $(x^1, .., x^k)$ be a local coordinate system centered
  at $p$. $p$ is called a non-degenerate critical point of $f$ if
  \begin{equation}
    \left( \frac{\partial^2 f}{\partial x^i \partial x^j} (p) \right)
  \end{equation}
  is an invertible matrix. On the other hand, $p$ is a called degenerate
  critical point if the matrix is singular.
\end{defn}

The definition above is independent of local coordinates. For, if $(x^1, ..,
x^k)$ and $(y^1, .., y^k)$ are two local coordinates at $p$, then
\begin{equation}
      \left( \frac{\partial^2 f}{\partial y^k \partial y^l} (p) \right) =
      J^T\left( \frac{\partial^2 f}{\partial x^i \partial x^j} (p) \right)J,
\end{equation}
where $J = \left( \frac{\partial x^i}{\partial y^j}(p) \right)_{i,j}$ is the
coordinate change matrix, which is non-singular.

The non-degeneracy of critical points also has an intrinsic meaning. Let $p$ be
a critical point. We define a symmetric bilinear form $f_{**}$ on $T_PM$, called
the \emph{Hessian} of $f$ at $p$. If $v, w \in T_pM$, then $v$ and $w$ have
extensions $\tilde{v}, \tilde{w}$. Define $f_{**}(v, w) =
\tilde{v}_p((\tilde{w})f)$. We must show that this is symmetric and
well-defined.$f_{**}$ is symmetric, since $\tilde{v}_p((\tilde{w})f) -
\tilde{w}_p((\tilde{v})f) = [\tilde{v}, \tilde{w}_p](f) = 0$\footnote{The last
  equality holds because $p$ is a critical point of $f$.} Therefore $f_{**}(v,
w) = v((\tilde{w})f) = w((\tilde{v})f)$. From both expressions of $f_{**}$ we
see that the definition is independent of the extensions of $v$ and $w$.

If $(x^1, .., x^n)$ is a local coordinate system, then the bilinear form
$f_{**}$ can be expressed as a symmetric matrix $\left(\frac{\partial^2
    f}{\partial x^i \partial x^j}(p)\right)$. Thus $p$ is a non-degenerate
critical point if the Hessian of $f$ at $p$ is a non-degenerate symmetric form.

Let $p$ be a critical point. We define the \emph{index of $f$ at $p$} as the
maximal dimension of subspaces in which $f_{**}$ is negative definite. The
\emph{nullity} is the subspace $V$ of $T_pM$ such that $f_{**}(V,w) = 0$ for all
$w \in T_pM$. The lemma of Morse shows that the behavior of $f$ at $p$ can be
completely described by this index. Before proving this, there is a small lemma:

\begin{lem}
  Let $f$ be a smooth function in a convex neighborhood $V$ of $0$ in $\RR^n$,
  with $f(0) = 0$. Then
  \begin{equation}
    f(x_1, .., x_n) = \sum_{i=1}^{n} x_ig_i(x_1, .., x_n)
  \end{equation}
  for some smooth $g_i$ defined in $V$, with $g_i(0) = \frac{\partial
    f}{\partial x_i}(0)$.
\end{lem}

\begin{proof}
  \begin{equation}
    \begin{aligned}
      f(x_1, .., x_n) &= \int_0^1 \frac{d f(tx_1, .., tx_n)}{dt} dt \\
                     &= \int_0^1 \sum_{i=1}^{n} x_i\frac{\partial f}{\partial
                       x_i}(tx_i) dt
    \end{aligned}
  \end{equation}
  Set $g_i = \int_0^1 \sum_{i=1}^{n} \frac{\partial f}{\partial x_i}(tx_i) dt$.
\end{proof}

\begin{lem}[Lemma of Morse]
  Let $p$ be a non-degenerate critical point for $f$. Then there is a local
  coordinate system $(y^1, .., y^n)$ in a neighborhood $U$ of $p$ with $y^i(p) =
  0$ and for all $i$ and such that the identity
  \begin{equation}
    f = f(p) - (y^1)^2 - .. - (y^\lambda)^2 + (y^{\lambda+1})^2 + .. + (y^n)^2
  \end{equation}
  where $\lambda$ is the index of $f$ at $p$.
\end{lem}

\begin{proof}
  Suppose such a local coordinate system exists. We prove that $\lambda$ must be
  the index of $f$ at $p$. If
  \begin{equation}
        f(q) = f(p) - (y^1(q))^2 - .. - (y^\lambda(q))^2 + (y^{\lambda+1}(q))^2 + .. + (y^n(q))^2,
  \end{equation}
  then
  \begin{equation}
    \frac{\partial^2 f}{\partial y^i \partial y^j} = 
    \begin{cases}
      -2 & i = j \le \lambda, \\
      2  & i = j > \lambda, \\
      0  & otherwise,
    \end{cases}
  \end{equation}
  which shows that $\lambda$ is the index of $f$ at $p$.

  We now show that such a local coordinate system exists.

  Let $(x^1, .., x^n)$ be a local coordinate system in a neighborhood $U$ of
  $p$, such that $x^i(p) = 0$ for all $i$. We may also suppose that $f(p) = 0$.
  By the previous lemma we can write
  \begin{equation}
    f(x_1, .., x_n) = \sum_j x_jg_j(x_1, .., x_n)
  \end{equation}
  Since $p$ is a critical point, 
  \begin{equation}
    0 = \frac{\partial f}{\partial x^j}(0) = g_j(0).
  \end{equation}

  Applying the lemma to $g_j$, we have
  \begin{equation}
    g_j(x) = \sum_i x_ih_{ij}(x_1, .., x_n).
  \end{equation}
  It follows that
  \begin{equation}
    f(x_1, .., x_n) = \sum_{i,j} x^ix^jh_{ij}(x_1, .., x_n).
  \end{equation}

  We may assume $h_{ij}$ is symmetric, since we can write $\bar{h}_{ij} =
  \frac{1}{2}(h_{ij}+h_{ji})$ and get $f = \sum_{i,j}
  x^ix^j\bar{h}_{ij}$. Moreover, $\bar{h}_{ij}(0) = \frac{1}{2}\frac{\partial^2
    f}{\partial x^i \partial x^j}(0)$, so it is non-singular.

  By induction, we proceed to find the local coordinate system to express $f$ in
  the desired form. Suppose by induction we have coordinates $u_1, .., u_n$ in a
  neighborhood $U$ of $p$ such that
  \begin{equation}
    f = \pm (u_1)^2 \pm .. \pm (u_{r-1})^2 + \sum_{i, j \ge r} u_iu_jH_{ij}(u_1,
    .., u_n),
  \end{equation}
  where $H_{ij}$ is symmetric and non-singular at $0$. After a linear
  change\footnote{If some of the diagonal elements is nonzero, say $H_{qq}$, we
    swap the two coordinates $u_r$ and $u_q$. If all diagonal elements are
    zeros, then we choose $k > r$ such that $\frac{\partial^2 f}{\partial
      u^r \partial u^{k}} \ne 0$. Define new coordinates $(w^1, .., w^n) = (u^1,
    .., u^{r-1}, \frac{u^r+u^k}{2}, u^{r+1}, .., u^{k-1}, \frac{u^r-u^k}{2},
    u^{k+1}, .., u^n)$. Then $\frac{\partial^2 f}{\partial w^r \partial w^r} =
    2\frac{\partial^2 f}{\partial u^r \partial u^k} \ne 0$, as desired.} in the
  last $n-r+1$ coordinates, we may assume that $H_{rr}(0) \ne 0$.Put $g(u_1, ..,
  u_n) = \sqrt{|{H_{rr}}|}$. Then there is a smaller neighborhood $U_2 \subset
  U_1$ of $p$ such that $g \ne 0$. A short calculation shows:

  \begin{equation}
    \begin{aligned}
      \sum_{i,j\ge r}u_iu_jH_{ij} 
      &= (u_r)^2H_{rr} + 2\sum_{i>r}u_iu_rH_{ir}+\sum_{i,j>r}u_iu_jH_{ij} \\
      &= \pm g(u_r+\sum_{i>r}\frac{u_iH_{ir}}{H_{rr}})^2+\sum_{i,j>r}u_iu_jH_{ij}',
    \end{aligned}
  \end{equation}
  where
  \begin{equation}
    H_{ij}' = H_{ij}-\frac{H_iH_j}{H_{rr}}
  \end{equation}
  is symmetric.

  Define new local coordinates
  $v_1$, .., $v_n$ by
  \begin{equation}
    v_i = 
    \begin{cases}
      u_i & i \ne r \\
      g(u_1, .., u_n)\left[u_r+\sum_{i>r}\frac{u_iH_{ir}}{H_{rr}}\right] & i = r
    \end{cases}
  \end{equation}

  By induction, the proof is complete.

\end{proof}

\begin{defn}
  A 1-parameter group of diffeomorphisms of a manifold $M$ is a smooth map
  \begin{equation}
    \phi : \RR \times M \to M
  \end{equation}
  such that
  \begin{itemize}
  \item for each $t \in \RR$ the map $\phi_t : M \to M$ defined by $\phi_t(q) =
    \phi(t, q)$ is a diffeomorphism of $M$ onto itself, and that
  \item for all $t,s \in \RR$, $\phi_{t+s} = \phi_t \circ \phi_s$.
  \end{itemize}
\end{defn}

\begin{lem}
  A smooth vector field on $M$ which vanishes outside of a compact set $K
  \subset M$ generates a unique 1-parameter group of diffeomorphisms on $M$.
\end{lem}

\begin{proof}
  Let $p \in M$. Then there exists a neighborhood $U$ of $p$ and $\epsilon > 0$
  such that the local flow $\phi : (-\epsilon, \epsilon)\times U \to M$ exists
  and is unique and the dependence on initial values is smooth. Since $K$ is
  compact, it can be covered by a finite number of $U$. Let $\epsilon_0$ be the
  smallest $\epsilon$ and set $\phi_t(q) = q$ for all $q \notin K$. Then the
  flow $\phi: (-\epsilon_0, \epsilon_0)\times M \to M$ exists and is unique. If
  $|s|,|t,|s+t| < \epsilon_0$ and set $f(t) = \phi_t(q)$. Then $\frac{df}{dt} =
  X_{f(t)}$ and $f(0) = \phi_s(q)$. Then $f(t) = \phi_{t+s}(q)$ by
  uniqueness. Thus $\phi_{t+s} = \phi_t \circ \phi_s$.

  We can extend the time domain from $(-\epsilon_0, \epsilon_0)$ to $\RR$ by
  define
  \begin{equation}
    \phi_t = \phi_{\frac{\epsilon_0}{2}} \circ \phi_{\frac{\epsilon_0}{2}} \circ
    .. \circ \phi_{\frac{\epsilon_0}{2}} \circ \phi_r,
  \end{equation}
  where $t = k\frac{\epsilon_0}{2}+ r$ and $|r| \le \frac{\epsilon_0}{2}$.
\end{proof}

\section{Homotopy Type in Terms of Critical Values}

Let $M$ be a smooth manifold and let $f: M \to \RR$ be a smooth function without
degenerate critical points. First we define
\begin{equation}
  M^a = f^{-1}(-\infty, a] = \{ p \in M | f(p) \le a \}.
\end{equation}

Then we investigate the ``topological difference'' between $M^a$ and $M^b$,
where $a < b$. The first case is simpler, namely there is no critical point in
$f^{-1}[a,b]$:

\begin{thm}
  \label{3.1}
  Suppose $f^{-1}[a,b]$ is compact and contains no critical points of $f$. Then
  $M^a$ is diffeomorphic to $M^b$. Furthermore, $M^a$ is a deformation retract
  of $M^b$, so that the inclusion map $\iota : M^a \to M^b$ is a homotopy
  equivalence.
\end{thm}

\begin{proof}
  Intuitively, we try to ``squeeze'' $M^b$ into $M^a$ in the direction of the
  gradient of $f$. Choose a Riemannian metric on $M$ and denote the inner
  product as $\langle X, Y \rangle$. The \emph{gradient} of $f$ is characterized
  by the identity\footnote{In fact, it is uniquely defined. To see this, choose
    local coordinates $u^1$, .., $u^n$ and put $X = \frac{\partial}{\partial
      u^i}$ into the identity. Write $grad \: f = f^j\frac{\partial}{\partial
      u^j}$. Then identity becomes $g_{ij}\frac{\partial}{\partial u^j} =
    \frac{\partial f}{\partial u^i}$. Taking inverse of the matrix $(g_{ij})$,
    we get $a^j = g^{ji}\frac{\partial f}{\partial u^i}$.}
  \begin{equation}
    \langle X, grad \: f\rangle = X(f)
  \end{equation}

  Let $\rho : M \to \RR$ be a smooth function which is equal to $1/\langle grad
  \: f, grad \: f \rangle$ throughout the compact set
  $f^{-1}[a,b]$\footnote{$grad \: f \ne 0$ in $f^{-1}[a,b]$; otherwise there is
    a critical point. It is easy to check this in terms of normal coordinates.}
  , and which vanishes outside of a compact neighborhood of this set. Then the
  vector field 
  \begin{equation}
    X = \rho (grad \: f)
  \end{equation}
  satisfies the condition in the last lemma of the previous section. Hence $X$
  generates a 1-parameter group of diffeomorphisms
  \begin{equation}
    \phi_t : M \to M.
  \end{equation}

  For fixed $q \in M$ consider the function $t \mapsto f(\phi_t(q))$. Then If
  $\phi_t(q) \in f^{-1}[a,b]$, then
  \begin{equation}
    \frac{d f(\phi_t(q))}{dt} = X_{\phi_t(q)}(f) = 
    \langle \rho(grad \:  f), grad \: f \rangle \le 1.
  \end{equation}
  Hence the inequality $f(\phi_t(q)) \le t + f(q)$ holds for all $t \in \RR$. It

  Consider the diffeomorphism $\phi_{b-a} : M \to M$. If $p \in M^b$, let $q =
  \phi_{a-b}(p)$. Then $\phi_{b-a}(q) = p$ and $f(q) = f(\phi_{a-b}(p)) \le
  a-b+f(p) \le a$. So $\phi_{b-a}$ maps $M^a$ onto $M^b$. It follows that
  $\phi_{b-a}$ maps $M^a$ diffeomorphically onto $M^b$. This prove the first
  part of the theorem.

  Now we construct a 1-parameter family of maps by
  \begin{equation}
    r_t(q) =
    \begin{cases}
      q & if f(q) \le a \\
      \phi_{t(a-f(q))}(q) & if a \le f(q) \le b.
    \end{cases}
  \end{equation}

  Then $r : [0,1] \times M^b \to M^a$ is a homotopy between identity of $M^a$
  and the deformation retract from $M^b$ to $M^a$. This completes the proof.
\end{proof}

The case when $f^{-1}[a,b]$ contains a critical point of $f$ is the essential
part of the Morse theory. It can be shown that $M^b$ is just $M^a$ with a number
of cells attached.

\begin{thm}
  \label{3.2}
  Let $f: M \to \RR$ be a smooth function and let $p$ be a non-degenerate
  critical point with index $\lambda$. Setting $f(p) = c$, suppose that $p$ is
  the only critical point in $f^{-1}[c-\epsilon, c+\epsilon]$, for some
  $\epsilon > 0$. Then for all $\epsilon$ sufficiently small, the set
  $M^{c-\epsilon}$ has the same homotopy type of $M^{c+\epsilon}$ with a
  $\lambda$-cell attached.
\end{thm}

The idea of the proof is to define a new smooth function $F: M \to \RR$ such
that inside a small neighborhood $H$ of $p$, $F < f$, and that $F = f$ outside
$H$. So $F^{-1}[-\infty, c-\epsilon] = M^{c-\epsilon} \cup H$ and
$F^{-1}[-\infty,c+\epsilon] = M^{c+\epsilon}$. If $F^{-1}[c-\epsilon,
c+\epsilon]$ has no critical point and is compact, then $M^{c-\epsilon} \cup H$
is a deformation retract of $M^{c+\epsilon}$ by the preceding theorem. If $H$
can be deformed into a $\lambda$-cell in it, the theorem is proved.

By the idea above, we divide the proof of the theorem into the following
steps.

Choose a coordinate system $u^1$, .., $u^n$ in a neighborhood $U$ of $p$ so that
the identity
\begin{equation}
  f = c - (u^1)^2 - .. - (u^{\lambda})^2 + (u^{\lambda+1})^2 + .. + (u^{n})^2
\end{equation}
holds throughout $U$ (Morse lemma). Then at $p$ we have $u^i(p) = 0$, $i = 1,
.., n$.

Choose $\epsilon > 0$ sufficiently small so that
\begin{itemize}
\item The region $f^{-1}[c-\epsilon,c+\epsilon]$ is compact and contains no
  critical points other than p.
\item The diffeomorphic image of $U$ in $\RR^n$ contains the closed ball
  $\{\sum_i (u^i)^2 \le 2\epsilon\}$.
\end{itemize}

Then $U$ contains a $\lambda$-cell, namely the inverse image of
$\{(u^1)^2+..+(u^{\lambda})^2 \le \epsilon, u^{\lambda+1}=..=u^{n} = 0\}$. Note
that $e^\lambda \cap M^{c-\epsilon}$ is exactly $\{f = c- \epsilon\} =
\{(u^1)^2+..+(u^\lambda)^2 = \epsilon \}$. So that the $\lambda$-cell is really
attached to $M^{c-\epsilon}$.

Next we construct a new smooth function $F : M \to \RR$ as follows. Let $\mu
:\RR \to \RR$ be a smooth function satisfying the conditions
\begin{equation}
  \begin{cases}
    \mu(0) > \epsilon \\
    \mu(r) = 0 \: \forall r \ge 2\epsilon \\
    -1 < \mu'(r) \le 0 \: \forall r.
  \end{cases}
\end{equation}

For example, we can define $\mu$ as follows. Let $\mu = \epsilon'
\frac{\int_x^{2\epsilon} \alpha(t)\alpha(2\epsilon-t)dt}{\int_0^{2\epsilon}
  \alpha(t)\alpha(2\epsilon-t)}dt$, where $2\epsilon > \epsilon' > \epsilon$ and
\begin{equation}
  \alpha(x) =
  \begin{cases}
    0 & x \le 0 \\
    e^{-1/x} & x > 0
  \end{cases}
\end{equation}
is smooth. Then $0 \ge \mu'(r) > -\frac{\epsilon'}{2\epsilon} > -1$.

Now let F coincides with f outside of the coordinate neighborhood $U$, and let
\begin{equation}
  F = f -\mu((u^1)^2+..+(u^{\lambda})^2 + 2(u^{\lambda+1})^2+..+(u^{n})^2)
\end{equation}
within $U$. It is easily verified that $F$ is a well-defined smooth function
throughout $M$.

It is convenient to write $\xi = (u^1)^2+..+(u^{\lambda})^2$ and $\eta =
(u^{\lambda+1})^2+..+(u^{n})^2$.

\begin{asser}
  The region $F^{-1}(-\infty, c+\epsilon]$ coincides with the region
  $M^{c+\epsilon}$.
\end{asser}

\begin{proof}
  Let $A = f^{-1}[-\infty, c+\epsilon]$, and let $B = F^{-1}[-\infty,
  c+\epsilon]$. If $q \in A$, then $F(q) \le f(q) \le c+\epsilon$ and so $q \in
  B$. To prove the converse, let $q \in B$. If $q \in \{ \xi + 2\eta \le
  2\epsilon \}$, then $f(q) = c - \xi + \eta \le c+ \frac{1}{2}\xi + \eta \le
  c+\epsilon$. If $q \notin \{ \xi + 2\eta \le 2\epsilon \}$, then $f(q) =
  F(q) \le c+\epsilon$. In both cases, $q \in A$.
\end{proof}

\begin{asser}
  The critical points of $F$ are the same as those of $f$.
\end{asser}

\begin{proof}
  \begin{equation}
    \nonumber
    \begin{aligned}
      F &= c - \xi + \eta - \mu(\xi+2\eta) \\
      \frac{\partial F}{\partial \xi} &= -1 - \mu'(\xi+2\eta) < 0 \\
      \frac{\partial F}{\partial \eta} &= 1 - 2\mu'(\xi+2\eta) \ge 1 \\
      dF &= \frac{\partial F}{\partial \xi}d\xi+\frac{\partial F}{\partial
        \eta}d\eta
    \end{aligned}
  \end{equation}
  Since the 1-forms $d\xi$ and $d\eta$ are simultaneously 0 only when both
  $\xi$ and $\eta$ equal 0, it follows that $F$ has no critical point in $U$
  other than $p$.
\end{proof}

Now, since $F \le f$, if $q \in F^{-1}[c-\epsilon, c+\epsilon]$, then
$c-\epsilon \le f(q)$. But $q \in F^{-1}(-\infty, c+\epsilon] = f^{-1}(-\infty,
c+\epsilon]$, $f(q) \le c+ \epsilon$. Therefore $q \in f^{-1}(-\infty,
c+\epsilon]$. This proves that $F^{-1}[c-\epsilon, c+\epsilon] \subset
f^{-1}[c-\epsilon, c+\epsilon]$. Being a closed subset of a compact set in a
locally compact Hausdorff space, $F^{-1}[c-\epsilon, c+\epsilon]$ is compact.
Note that $F(p) = c- \mu(0) < c-\epsilon$. $p \notin F^{-1}[c-\epsilon,
c+\epsilon]$. Together with the preceding theorem we get

\begin{asser}
  $F^{-1}(-\infty, c-\epsilon]$ is a deformation retract of $F^{-1}[-\infty,
  c+\epsilon]$. \qed
\end{asser}

Write $F^{-1}(-\infty, c-\epsilon]$ as $M^{c-\epsilon}\cup H$, where $H$ is the
closure of $F^{-1}(-\infty, c-\epsilon] - M^{c-\epsilon}$. In fact, $H = \{
\epsilon-\mu(\xi+2\eta) \le \xi-\eta \le \epsilon \}$. Define
\begin{equation}
  e^\lambda = \{ q | \xi(q) \le \epsilon, \eta(q) = 0\}
\end{equation}
Then $e^\lambda \subset H$. For, if $q \in e^\lambda$, then $f(q) >
c-\epsilon$. Thus $q \notin M^{c-\epsilon}$. Also, since $\frac{\partial
  F}{\partial \xi} < 0$, $F(q) < F(p) < c-\epsilon$. Thus $q \in F^{-1}(-\infty,
c-\epsilon]$.

\begin{asser}
  $M^{c-\epsilon} \cup e^\lambda$ is a deformation retract of
  $M^{c-\epsilon}\cup H$.
\end{asser}

\begin{proof}
  We operate on the diffeomorphic image on $\RR^n$. In the $\xi-\eta$ plane, $H$
  is the region $F \ge c-\epsilon$ minus the region $f \ge c-\epsilon$.

  Denote the homotopy between the retraction and identity by $r_t : [0,1]\times
  (M^{c-\epsilon}\cup H) \to M^{c-\epsilon} \cup e^\lambda$.

  Decompose $H$ into 2 regions: (1) $\xi \le \epsilon$ and (2) $\epsilon \le \xi
  \le \eta + \epsilon$.

  In region (1), we simply project $H$ into $e^\lambda$. The homotopy is thus
  given by 
  \begin{equation}
    r_t: (t, (u^1,..,u^n)) \mapsto (u^1, .., u^\lambda, tu^{\lambda+1}, .., tu^n).
  \end{equation}

  We check that the homotopy lies entirely in $H =
  \{\epsilon-\mu(\xi+2\eta) \le \xi-\eta \le \epsilon \}$. It does,
  indeed, since $\epsilon \ge \xi \ge \xi-t^2\eta \ge \xi-\eta \ge
  \epsilon-\mu(\xi+2\eta) \ge \epsilon-\mu(\xi+2t^2\eta)$.

  In region (2), we project this part into $M^{c-\epsilon}$, in the
  $\eta$-direction as the in case (1). Let the following denote the projection:
  \begin{equation}
    (u^1, .., u^n) \mapsto
    (u^1, .., u^\lambda, ku^{\lambda+1}, .., ku^n) \in M^{c-\epsilon}
  \end{equation}

  Then it's easy to calculate $k =
  (\frac{\xi-\epsilon}{\eta})^{\frac{1}{2}}$. Then the homotopy is simply the
  line along the projection:
  \begin{equation}
    r_t: (t, (u^1,..,u^n)) \mapsto
    (u^1, .., u^\lambda, s_tu^{\lambda+1}, .., s_tu^n),
  \end{equation}
  where
  \begin{equation}
    s_t = t + (1-t)k.
  \end{equation}

  Again we check that the homotopy lies entirely in $H$. We first note that $s_t
  \in [0,1]$. Then $\xi-s_t^2\eta \ge \xi-\eta \ge \epsilon-\mu$. Also,

  \begin{equation}
    \begin{aligned}
      \xi-s_t^2\eta &= \xi -
      (t^2+2t(1-t)(\frac{\xi-\epsilon}{\eta})^{1/2}+(1-t)^2\frac{\xi-\epsilon}{\eta})\eta
      \\
      &=
      \xi-t^2\eta-2(t-t^2)\sqrt{\eta(\xi-\epsilon)}-(1-t)^2\epsilon+(1-t)^2\epsilon
      \\
      &\le
      \xi-t^2(\epsilon-\xi)-(2t-2t^2)(\xi-\epsilon)-(1-t)^2\xi+(1-t)^2\epsilon
      \\
      &= \epsilon,
    \end{aligned}
  \end{equation}
  where we use the condition $\xi-\eta \le \epsilon$.

  Since $s_t$ is continuous and it is compatible with the case (1) when $\xi =
  \epsilon$, $r_t$ is well-defined and continuous on the union.

  It remains to define $r_t$ on $M^{c-\epsilon}$. We simply define it as the
  identity map. Note that this definition is compatible with that of case (2),
  since on $\xi-\eta=\epsilon$, $s_t = 1$.

  The assertion follows.
\end{proof}

As the last assertion is proved, so is the theorem.

\begin{rmk}
  \label{3.3}
  More generally suppose that there are $k$ non-degenerate critical points
  $p_1$, .., $p_k$ with indices $\lambda_1$, .., $\lambda_k$ in
  $f^{-1}(c)$. Then a similar proof shows that $M^{c+\epsilon}$ has the homotopy
  type of $M^{c-\epsilon}\cup e^{\lambda_1}\cup .. \cup e^{\lambda_k}$.
\end{rmk}

\begin{thm}
  \label{main}
  If f is a differentiable function on a manifold $M$ with no degenerate
  critical points, and if each $M^a$ is compact, then $M$ has the homotopy type
  of a CW-complex with one cell of dimension $\lambda$ for each critical point
  of index $\lambda$.
\end{thm}

The proof will be based on two lemmas concerning a topological space $X$ with a
cell attached. The first lemma says that if two attaching methods are homotopic
to each other, then the two resulting spaces are homotopy equivalent. The second
lemma shows that if the base space is changed by a map preserving its homotopy
type, then then the two attaching spaces are homotopic equivalent in a natural
way.

\begin{lem}[Whitehead]
  \label{whitehead}
  Let $\phi_0$ and $\phi_1$ be homotopic maps from the sphere $\dot{e}^\lambda$
  to $X$. Then the identity map of $X$ extends to a homotopy equivalence
  \begin{equation}
    k: X \cup_{\phi_0} e^\lambda \to X \cup_{\phi_1} e^\lambda
  \end{equation}
\end{lem}

\begin{proof}
  Let $\phi_t$ be the homotopy between $\phi_0$ and $\phi_1$. Let $u$ denote a
  point on $\dot{e}^\lambda$. We define $k$ by the following formulas:
  \begin{equation}
    \begin{array}{lcl}
      k(x) = x, & x \in X, \\
      k(tu) = 2tu & t \in [0,\frac{1}{2}], \\
      k(tu) = \phi_{2-2t}(u) & t \in [\frac{1}{2},1].
    \end{array}
  \end{equation}

  Notice that $k(\frac{1}{2}u) = u = \phi_1(u)$.

  Define $l : X \cup_{\phi_1} e^\lambda \to X \cup_{\phi_0} e^\lambda$ in a
  similar way as $k$:

  \begin{equation}
    \begin{array}{lcl}
      k(x) = x, & x \in X, \\
      k(tu) = 2tu & t \in [0,\frac{1}{2}], \\
      k(tu) = \phi_{2t-1}(u) & t \in [\frac{1}{2},1].
    \end{array}
  \end{equation}

  Then $lk$ is defined as the following formulas

  \begin{equation}
    \begin{array}{lcl}
      lk(x) = x, & x \in X, \\
      k(tu) = 4tu & t \in [0,\frac{1}{4}], \\
      k(tu) = \phi_{4t-1}(u) & t \in [\frac{1}{4},\frac{1}{2}], \\
      k(tu) = \phi_{2-2t}(u) & t \in [\frac{1}{1},1].
    \end{array}
  \end{equation}

  Hence there is a homotopy between $lk$ and identity

  \begin{equation}
    \begin{array}{lcl}
      H(s, x) = x, & s \in [0,1], x \in X,\\
      H(s, tu) = (4-3s)tu & t \in [0,\frac{1}{4-3s}], \\
      H(s, tu) = \phi_{4t-1}(u) & t \in [\frac{1}{4-3s},\frac{2-s}{4-3s}], \\
      H(s, tu) = \phi_{\frac{4-3s}{4}(2-2t)}(u) & t \in [\frac{2-s}{4-3s},1].
    \end{array}
  \end{equation}

  $H$ is continuous on a finite number of closed sets of $[0,1]\times[0,1]$ and
  well-defined, so $H$ is continuous.

  In a similar fashion, we can construct a homotopy from $kl$ to identity

  \begin{equation}
    \begin{array}{lcl}
      \bar{H}(s, x) = x, & s \in [0,1], x \in X,\\
      \bar{H}(s, tu) = (4-3s)tu & t \in [0,\frac{1}{4-3s}], \\
      \bar{H}(s, tu) = \phi_{2-(4-3s)t}(u) &
      t \in [\frac{1}{4-3s},\frac{2-s}{4-3s}], \\
      \bar{H}(s, tu) = \phi_{\frac{(4-3s)t}{2}+\frac{3}{2}s-1}(u) &
      t \in [\frac{2-s}{4-3s},1].
    \end{array}
  \end{equation}

\end{proof}

\begin{lem}
  \label{3.7}
  Let $\phi : \dot{e}^\lambda \to X$ be an attaching map. Any homotopy
  equivalence $f: X \to Y$ extends to a homotopy equivalence
  \begin{equation}
    F: X \cup_\phi e^\lambda \to Y \cup_{f\phi} e^\lambda
  \end{equation}
\end{lem}

\begin{proof}
  Define $F|_X = f$, $F|_{e^\lambda} = identity$. Let $g$ be the homotopy
  inverse of $f$. Define $G: Y \cup_{f\phi} e^\lambda \to X \cup_{gf\phi}
  e^\lambda$ by $G|_Y = g$, $G|_{e^\lambda} = identity$. Since $gf\phi$ is
  homotopic to $\phi$, by \ref{whitehead} there is a homotopy equivalence
  $G: X \cup_{gf\phi} e^\lambda \to X \cup_{\phi} e^\lambda$. We want to show
  that $kG$ is a homotopic inverse of $F$. We first write down the composition
  $kGF$ in the following formulas

  \begin{equation}
    \begin{array}{lcl}
      kGF(x) = x, & x \in X, \\
      kGF(tu) = 2tu & t \in [0,\frac{1}{2}], \\
      kGF(tu) = h_{2-2t}(u) & t \in [\frac{1}{2},1]
    \end{array}
  \end{equation}
  where $h_t$ is a homotopy between $gf$ and identity. $kGF$ is homotopic to
  identity by the following formulas

  \begin{equation}
    \begin{array}{lcl}
      H(s, x) = h_s(x), & s \in [0,1], x \in X,\\
      H(s, tu) = (2-s)tu & t \in [0,\frac{1}{2}], \\
      H(s, tu) = h_{2-(2-s)t}(tu) & t \in [\frac{1}{2},1]
    \end{array}
  \end{equation}

  Thus $F$ has a left inverse. Note that $g$ and the induced $G$ also satisfy
  the condition of the lemma. So $G$ also has a left inverse.

  It is elementary that if $F$ has both a left inverse $L$ and a right inverse
  $R$, then $F$ is a homotopy equivalence and both $L$ and $R$ are two sided
  inverses.

  Now we have $k(GF) \sim id$. Since $k$ is a homotopy equivalence, $GFk \sim
  id$. Thus $G$ has a right inverse. Then $G$ is also a homotopy equivalence and
  $FkG \sim id$. Thus $F$ also has a right inverse. This proves that $F$ is a
  homotopy equivalence.
\end{proof}

We now go back to the proof of theorem \ref{main}:

\begin{proof}
  Let $c_1 < c_2 < c_3 < ..$ be the critical values of $f: M \to \RR$. The
  sequence ${c_i}$ has no cluster points, for each $M^a$ is compact \footnote{If
    $c$ is a cluster point of ${c_i}$,then there exists a sequence ${p_j}$ in M
    such that $f(p_j) = c_{i_j}$ and $\lim_{j \to \infty} p_j = p$ If p is not a
    critical point, then there is a neighborhood $U$ of $p$ such that each point
    of $U$ is not a critical point. But this is a contradiction, since $p$ is
    the limit of the sequence ${p_j}$.}. We first prove by induction that each
  level set $M^a$ has the homotopy type of a CW-complex. If $a < c_1$, then $M^a
  = \emptyset$. Suppose $c_{i-1} < a < c_{i} =: c$. By theorems \ref{3.1},
  \ref{3.2} and \ref{3.3}, $M^{c+\epsilon} = M^a \cup_{\phi_1} e^{\lambda_1}
  \cup .. \cup_{\phi_{j(c)}} e^{\lambda_{j(c)}}$, for sufficiently small
  $\epsilon$. By induction hypothesis, there is a homotopy equivalence $h : \to
  K$.By lemma \ref{3.7}, $h$ induces a homotopy equivalence between
  $M^{c+\epsilon} = M^a \cup_{\phi_1} e^{\lambda_1} \cup .. \cup_{\phi_{j(c)}}
  e^{\lambda_{j(c)}}$ and $K \cup_{h\circ\phi_1} e^{\lambda_1} \cup
  .. \cup_{\circ\phi_{j(c)}} e^{\lambda_{j(c)}}$

  Each $h \circ \phi_j$ is homotopic by cellular approximation to
  \begin{equation}
    \psi_j: \dot{e}^{\lambda_j} \to K^{\lambda_j-1}.
  \end{equation}

  By lemma \ref{whitehead}, replacing the attaching map does not
  change the homotopy type. Thus we have shown that $M^{c+\epsilon}$
  has the homotopy type of a CW-complex.

  If M is compact, then $M = M^a$ for some $a \in \RR$ and the proof is done. If
  M is not compact, but all critical points lie in $M^a$, a proof similar to
  \ref{3.1} shows that $M^a$ is a deformation retract of $M$.

  If there are infinitely many critical points, then we can construct an
  infinite sequence of homotopy equivalences
  \begin{equation*}
    \xymatrix@C+0em@R+0em{
      M_{a_1} \ar[d]_{g_1} \ar@{^(->}[r] &
      M_{a_2} \ar[d]_{g_2} \ar@{^(->}[r]&\cdots \\
      K_1 \ar@{^(->}[r] & K_2 \ar@{^(->}[r] &\cdots
    }
  \end{equation*}
  each extending the previous one. Let $K = \cup_{i=1}^{\infty}K_i$ have the
  direct limit topology. We show that $K$ is a CW-complex. Let $C$ be a closed
  set in $K$. Then by definition, $C\cap K_i$ is closed in $K_i$ for all
  $i$. This means that $C\cap e_{ij}$ is closed in $e_{ij}$ for all $e_{ij} \in
  K_i$ for all $i$. Thus $C \cap e$ is closed for all cell in $K$. This proves
  that $K$ has the weak topology.

  Now let $C$ be a compact set in $K$. Let $I=\{ p_i | p_i \in C \cap
  e_i\}$, where $e_i$ ranges over all cells in $K$. Then $I$ is
  closed, since the intersection of $I$ and each cell is a point. $I$
  is descrete, since each $p \in I$ lies in a cell and each cell is
  closure-finite. This proves that $K$ has the closure-finite
  property.

  Let $g: M \to K$ be the limit map. We need Whitehead's theorem to conclude
  that $g$ is a homotopy equivalence:

  \begin{thm}
    If $M$ and $K$ are both dominated by CW-complexes, then any map $M \to K$
    which induces isomorphisms of homotopy groups is a homotopy equivalence.
  \end{thm}

  \begin{defn}\cite{Hat}
    A space $Y$ is dominated by a space $X$ if there are maps
    \begin{equation}
      \xymatrix{
        Y \ar[r]^i & X \ar[r]^r & Y
      }
    \end{equation}
    such that $ri \sim 1_Y$.
  \end{defn}

  $M$ can be dominated by a CW-complex in the following way. Embed $M$ into
  $\RR^n$. Let $N$ be a tubular neighborhood of $M$. Triangulate $\RR^n$ so that
  the simplex intersecting $M$ lies entirely in $N$. Collecting all the
  simplexes and their faces, we get a CW-complex $K$. $M \subset W$. The
  retraction of the tubular neighborhood into $M$ restricting on $W$ is a
  retraction on $W$ into $M$. This shows that $M$ is dominated by a CW-complex.

  It remains to show that $g_*$ is an isomorphism of each $\pi_n(M)$ and
  $\pi_n(K)$. We prove the onto-ness first. Let $f_i$ denote the homotopy
  inverse of $g_i$. Let $[\phi] \in \pi_n(K)$. Since $\phi(S^n)$ is compact,
  $\phi(S^n)$ lies in some $K_i$. Since $g_i\circ f_i \circ \phi$ is homotopic
  to $\phi$ in $K_i \subset K$, $g_*[f_i\circ\phi] = [\phi]$. Next we prove that
  $g_*$ is one-to-one. Suppose there exist $\phi_0, \phi_1 : S^n \to K$ such
  that $g\circ\phi_0 \sim g\circ\phi_1$ in $K$. Denote the homotopy as
  $H$. Since $H$ is continuous and $[0,1]\times S^n$ is compact, the homotopy
  lies in a compact set of $K$, thus in $K_i$ for some $i$. Therefore $f_i\circ
  g_i\circ\phi_0 \sim f_i\circ g_i\phi_1$ in $M^{a_i}$. Hence $f_i\circ
  \phi_0 \sim \phi_1$ in $M$.

  The theorem follows.

\end{proof}

\section{Manifolds in Euclidean Space: The Existence of Non-degenerate
  Functions}

Let $M$ be a smooth manifold of dimension $k$. By Whitney Embedding Theorem, we
can embed $M$ into $\RR^n, n > k$. Therefore, without loss of generality we may
assume $M \subset \RR^n$. For each $p \in \RR^n$, define a smooth function $L_p
: M \to \RR $ by
\begin{equation}
  L_p(q) = |q-p|^2 = (q-p) \cdot (q-p) = q\cdot q-2q \cdot p + p \cdot p
\end{equation}

We want to find a $p \in \RR^n$ such that $L_p$ has no degenerate critical
points in $M$. Denote $L_p$ as $f$. Let $q \in M$ and let $(u^1, .., u^k)
\mapsto (x^1(u), .., x^n(u))$ be a local coordinate system centered at $q$. Then $q$ is
a critical point if $\frac{\partial f}{\partial u^i}(q) = 0$ for all $i = 1, ..,
k.$; that is,
\begin{equation}
  \frac{\partial \vect{x}}{\partial u^i}(q) \cdot (q-p) = 0.
\end{equation}
It follows that $q \in M$ is a critical point of $L_p$ if and only if $p$ lies
in a line perpendicular to $M$ at $q$. Differentiate again, and we get
\begin{equation}
  \frac{\partial^2 f}{\partial u^i \partial u^j}(q) = 
  2\left(
    \frac{\partial^2 \vect{x}}{\partial u^i \partial u^j}(q) \cdot (q-p) +
    \frac{\partial \vect{x}}{\partial u^i}(q) \cdot \frac{\partial \vect{x}}{\partial u^j}(q)
  \right)
\end{equation}

We get the following conclusions:

$L_p$ is non-degenerate if and only if

\begin{itemize}
\item For each critical point $q$, $p$ lies in a line perpendicular to $M$ at
  $q$.
\item the matrix $\left(
    \frac{\partial^2 \vect{x}}{\partial u^i \partial u^j}(q) \cdot (q-p) +
    \frac{\partial \vect{x}}{\partial u^i}(q) \cdot \frac{\partial \vect{x}}{\partial u^j}(q)
  \right)$
  is non-singular.
\end{itemize}

Thus the main purpose of this section is to establish the existence of such $p
\in \RR^n$.

\subsection{Focal Points}

Let $N = \{ (q,v) \in \RR^{2n} | q \in M, v \in (T_qM)^{\perp} \subset \RR^n \}$
be the normal bundle of $M$ in $\RR^n$. Then $N$ is a manifold of dimension
$k+(n-k)=n$. We define a map $E$ that maps $N$ into the ambient space $\RR^n$ by
\begin{equation}
  E(q, v) = q + v.
\end{equation}
This is a differentiable map between two manifolds of dimension $n$. The point
$q+v$ thus lies in a normal line that is perpendicular to $M$ at $p$. This leads
to the definition of focal points:

\begin{defn}
  $e \in \RR^n$ is a focal point of $(M, q)$ with multiplicity $\mu > 0$ if $e =
  q+v$, where $(q, v) \in N$ and the Jacobian of $E$ at $(q, v)$ has nullity
  $\mu > 0$. $e$ is called a focal point of $M$ if $e$ is a focal point of $(M,
  q)$, $q \in M$.
\end{defn}

Next we fix a $\vect{q} \in M$ and a unit normal vector $\vect{v} \in \RR^n$ and
study the focal points of $(M, q)$ along the line $\vect{l} =
\vect{q}+t\vect{v}$. Before this, we first introduce the second the second
fundamental form of a manifold in $\RR^n$:

\begin{defn}
  Let $\vect{q} \in M$ and let $\vect{v} \in (T_qM)^\perp$, $|\vect{v}| =
  1$. When $g_{ij} = \delta_{ij}$, the second fundamental form\footnote{The more
    general definition is $-(\nabla \tilde{v})^T$, which is a linear
    transformation on $T_qM$. If $(u^1, .., u^k)$ is a normal coordinate system,
    the matrix of the second fundamental from is given by $(-\nabla^T
    \tilde{v})_{ij} = -\frac{\partial \tilde{v}}{\partial u^j} \cdot
    \frac{\partial x}{\partial u^i} = v \cdot \frac{\partial^2 x}{\partial
      u^i \partial u^j}$.} in the direction of $\vect{v}$ is defined as the
  matrix

  \begin{equation}
    \left(
      \vect{v} \cdot \vect{l}_{ij}
    \right)
  \end{equation}
  where $\vect{l}_{ij} = \frac{\partial^2 \vect{x}}{\partial u^i \partial
    u^j}(\vect{q})$.
  The eigenvalues of the matrix is called the principal curvatures $K_1$, ..,
  $K_k$ of $M$ at $\vect{q}$ in the normal direction $\vect{v}$.
\end{defn}

Now comes the key lemma in this section:

\begin{lem}
  The focal points of $(M, q)$ along $\vect{l}$  are precisely the points
  $\vect{q}+K_i^{-1}\vect{v}$, where $1 \le i \le k$, $K_i \ne 0$. Thus there
  are at most $k$ focal points of $(M,q)$ along $\vect{l}$, each being counted
  with its proper multiplicity.
\end{lem}

\begin{proof}
  First we are going to find a local coordinate system of $N$ centered at $(q,
  v)$. Let $\vect{w}_1(u^1, .., u^k)$, .., $\vect{w}_{n-k}(u^1, .., u^k)$ be
  orthonormal vector fields in $N$. Then we have a local coordinate system of
  $N$ at $(q, v)$ defined by
  \begin{equation}
    (u^1, .., u^k, t^1, .., t^{n-k}) \mapsto (\vect{x}(u),
    \sum_{\alpha=1}^{n-k}t^{\alpha} \vect{w}_{\alpha}(u^1, .., u^k)) \in N
  \end{equation}
  Write $E: N \to \RR$ in the coordinates as $E(u, t) = \vect{e}(u, t) \in
  \RR^n$ and get
  \begin{equation}
    \vect{e}(u, t) = \vect{x}(u) + \sum_{\alpha=1}^{n-k}t^{\alpha} \vect{w}_{\alpha}(u^1, .., u^k))
  \end{equation}
  with partial derivatives
  \begin{equation}
    \left\{
      \begin{array}{lcl}
        \frac{\partial \vect{e}}{\partial u^i} & = &
        \frac{\partial \vect{x}}{\partial u^i} + 
        \sum_{\alpha} t^{\alpha}\frac{\vect{w}_{\alpha}}{\partial u^i}  \\
        \frac{\partial \vect{e}}{\partial t^\beta} & = & \vect{w}_\beta
      \end{array}
    \right.
  \end{equation}

  Since the n-vectors $\frac{\partial \vect{x}}{\partial u^1}$, ..,
  $\frac{\partial \vect{x}}{\partial u^k}$, $\vect{w}_1$, .., $\vect{w}_{n-k}$
  form a basis of $T_q\RR^n$, the matrix
  \begin{equation}
    A = \left(
      \begin{matrix}
        \left(\frac{\partial \vect{x}}{\partial u^1}\right)^T \\
        \vdots \\
        \left(\frac{\partial \vect{x}}{\partial u^k}\right)^T \\
        \left(\vect{w}_1\right)^T \\
        \vdots \\
        \left(\vect{w}_{n-k}\right)^T
      \end{matrix}
    \right)
  \end{equation}
  is non-singular (we represent a vector as a column matrix). Multiply $A$ with
  the Jacobian $J$ of $E(u, t)$ and we get
  \begin{equation}
    \begin{aligned}
      AJ &= 
      \left(
        \begin{matrix}
          \left(\frac{\partial \vect{x}}{\partial u^1}\right)^T \\
          \vdots \\
          \left(\frac{\partial \vect{x}}{\partial u^k}\right)^T \\
          \left(\vect{w}_1\right)^T \\
            \vdots \\
          \left(\vect{w}_{n-k}\right)^T
        \end{matrix}
      \right)
      \left(
        \begin{matrix}
          \frac{\partial \vect{e}}{\partial u^1} &
          \cdots &
          \frac{\partial \vect{e}}{\partial u^k} &
          \frac{\partial \vect{e}}{\partial t^1} &
          \cdots &
          \frac{\partial \vect{e}}{\partial t^{n-k}} &
        \end{matrix}
      \right) \\
      &=
      \left(
        \begin{matrix}
          \left( \frac{\partial \vect{x}}{\partial u^i} \cdot \frac{\partial
              \vec{x}}{\partial u^j} + \sum_{\alpha}
            t^{\alpha}\frac{\vec{w}_{\alpha}}{\partial u^i} \cdot \frac{\partial
              \vect{x}}{\partial u^j} \right) & 0
          \\
          \sum_{\alpha} t^{\alpha}\frac{\vect{w}_{\alpha}}{\partial u^i} \cdot
          \vect{w}_{\beta} & \id
        \end{matrix}
      \right)
    \end{aligned}
  \end{equation}

  Since $J$ has the same rank of $AJ$, $J$ has the same nullity as the upper
  left block of $AJ$. Using the identity
  \begin{equation}
    \begin{aligned}
      0 &= \frac{\partial}{\partial u^i}\left(\vect{w}_{\alpha} \cdot
        \frac{\partial
          \vect{x}}{\partial u^j}\right) \\
      &= t^{\alpha}\frac{\vect{w}_{\alpha}}{\partial u^i} \cdot \frac{\partial
        \vect{x}}{\partial u^j} + \vect{w}_{\alpha} \cdot \frac{\partial^2
        \vect{x}}{\partial u^i \partial u^j},
    \end{aligned}
  \end{equation}
  we see that this upper left hand block is just the matrix 
  \begin{equation}
    \left( g_{ij} - \sum_{\alpha}t^{\alpha}\vect{w}_{\alpha} \cdot \vect{l}_{ij} \right)
  \end{equation}

  We choose the Riemman normal coordinates at $q$, so that $g_{ij}(q) =
  \delta_{ij}$. Now, $\sum_{\alpha}t^{\alpha}\vect{w}_{\alpha} = t\vect{v}$. The
  matrix
  \begin{equation}
    \label{eq:key}
    \left(
      \delta_{ij}-t\vect{v} \cdot  \vect{l}_{ij}
    \right)
  \end{equation}
  is singular if and only if $\frac{1}{t} = K_i$. Furthermore, $\mu$ equals the
  number of distinct $K_i$, each of which counted with its own
  multiplicity. This proves the lemma.
\end{proof}

Suppose $\vect{p} = \vect{q}+t\vect{v}$. Then $\vect{q}$ is a critical point of
$L_p$. Recall the criterion of $q$ being a non-degenerate critical point of
$L_p$. The criterion is exactly Eq (\ref{eq:key}). We thus obtain the
correspondence between the non-degeneracy of $L_p$ and the idea of focal points:

\begin{cor}
  The function $L_p$ is non-degenerate if and only if $p$ is not a focal point
  of $M$. \qed
\end{cor}

$L_p$ is degenerate if and only if $p$ is a focal point, which is a critical
value of a smooth map. By the following theorem of Sard, we see that there are
infinitely many $p \in \RR^n$ such that $L_p$ is non-degenerate:

\begin{thm}[Sard]
  Let $f: M^m \to N^n$ be a smooth map between two smooth manifolds. Then the
  critical values has measure 0 in $N^n$. \qed
\end{thm}

This established the existence of non-degenerate functions on $M$. As a result,
the main theorem in this report is proved:

\begin{thm}
  A differentiable manifold has the homotopy type of a CW-complex.\qed
\end{thm}

The preceding corollary can be sharpened as follows. Let $k \ge 0$ be an integer
and let $K \subset M$ be a compact set.

\begin{cor}
  Any bounded smooth function $f: M \to \RR$ can be uniformly approximated by a
  smooth function $g$ which has no degenerate critical points. Furthermore $g$
  can be chosen so that the i-th derivatives of $g$ on the compact set $K$
  uniformly approximate the corresponding derivatives of $f$, for $i \le k$.
\end{cor}

\begin{proof}
  Choose some embedding $h:M \to \RR^n$ of $M$ as a bounded subset of some
  euclidean space so that the first coordinate $h_1$ is exactly
  $f$. \footnote{By Whitney embedding theorem, there is an embedding $\phi : M
    \in R^n$. Compose each coordinate of $\phi$ with $arctan$, we get a bounded
    embedding, since $\psi = (\tan^{-1} \circ \phi^1, .., \tan^{-1} \circ
    \phi^n)$ is one-to-one and the differential map is injective. Let $h = (f,
    \tan^{-1}\phi)$. Then $h$ is an embedding from $M$ into $\RR^{n+1}$ such
    that $h_1 = f$.} Let c be a large number. Choose a point
  \begin{equation}
    p = (-c+\epsilon_1, \epsilon_2, .., \epsilon_n)
  \end{equation}
  close to $(-c, 0, .., 0) \in \RR^n$ such that the function $L_p: M \to \RR$ is
  non-degenerate. Define
  \begin{equation}
    g(x) = \frac{L_p(x)-c^2}{2c}
  \end{equation}
  A short computation gives
  \begin{equation}
    \begin{aligned}
      g(x) &= \frac{\sum_i h_i^2(x)-h_i(x)p_i+p_i^2-c^2}{2c} \\
           &= f(x) + \frac{\sum_i h_i^2(x)-\sum_i h_i(x)\epsilon_i+\sum_i \epsilon_i^2}{2c}-\epsilon_1.
    \end{aligned}
  \end{equation}
  Since $K$ is compact and each $h_i$ is continuous and bounded, it's easy to
  see that $g(x)$ approximates $f(x)$ uniformly by choosing $c$ large enough and
  $\epsilon_i$ small enough. 
\end{proof}

The preceding theory can also be used to describe the index of the function
$L_p$ at a critical point:

\begin{lem}[Index theorem for $L_p$]
  \label{index}
  \label{index_of_Lp}
  The index of $L_p$ at a non-degenerate critical point $q \in M$ is equal to
  the number of focal points of (M, q) which lie on the segment from $q$ to $p$;
  each focal point being counted with its multiplicity.
\end{lem}

\begin{proof}
  The index of the matrix 
  \begin{equation}
    \left( \frac{\partial^2 L_p}{\partial u^i \partial
        u^j}(q) \right) = 2(g_{ij}-t\vect{v} \cdot \vect{l}_{ij})
  \end{equation}
  equals the number of negative eigenvalues counted with multiplicity. Recall
  that we can choose the normal coordinates such that $(g_{ij})$ is
  identity. Then$\left( \frac{\partial^2 L_p}{\partial u^i \partial u^j}(q)
  \right)$ has a negative eigenvalue if and only if $1-K_it < 0$ for some
  principal curvatures $K_i$. Thus $K_i > \frac{1}{t}$. Recall that the focal
  points that lies on the line from $q$ through $p$ are of the form
  $\vect{q}+K_i^{-1}\vect{v}$. Therefore the correspondence established.

\end{proof}

\section{The Lefschetz Theorem on Hyperplane Sections}

As an application of the previous results, we will prove the Lefschetz theorem
on hyperplane sections in this section. Before the development, we should show
that every complex analytic manifold can be viewed as a CW-complex:

\begin{thm}
  A complex manifold analytic manifold $M$ of complex dimension $k$,
  bianalytically embedded as a closed subset of $\CC^n$ has the homotopy type of
  a $k$-dimensional CW-complex.
\end{thm}

The proof will be broken up into several steps. First we consider a quadratic
form in $k$ complex variables
\begin{equation}
  Q(z^1, .., z^k) = \sum b_{hj}z^hz^j.
\end{equation}

If we substitute $x^h+iy^h$ for $z^h$ and take the real part of $Q$, we obtain a
real quadratic form in $x^1, .., x^k, y^1, .., y^k$:

\begin{equation}
  Q'(x^1, .., x^k, y^1, .., y^k) = Re\{\sum b_{hj}z^hz^j\}.
\end{equation}

\begin{asser}
  If $e$ is an eigenvalue of $Q'$ with multiplicity $\mu$, then $-e$ is also an
  eigenvalue with the same multiplicity $\mu$.
\end{asser}

\begin{proof}
  First we note that $Q(iz^1, .., iz^k) = -Q(z^1, .., z^k)$. Therefore, in terms
  of $Q'$, we have
  \begin{equation}
    -Q' = J^TQ'J
  \end{equation}
  where
  \begin{equation}
    J = 
    \left(
      \begin{matrix}
        O & -I_k \\
        I_k & O
      \end{matrix}
    \right)
  \end{equation}
  is an orthogonal matrix. Therefore
  \begin{equation}
    \begin{aligned}
      det(eI-Q') = 0 &\Leftrightarrow det(eI-J^TQ'J) = 0 \\
      &\Leftrightarrow det(eI+Q') = 0.      
    \end{aligned}
  \end{equation}
\end{proof}

Now consider a complex manifold $M$ which is bianalytically embedded as a subset
of $\CC^n$. Let $q$ be a point of $M$.

\begin{asser}
  The focal points of $(M, q)$ along any normal lines occur in pairs
  which are situated symmetrically about $q$.
\end{asser}

\begin{proof}
  Let $(z^1, .., z^k)$ be coordinates centered at $q$ such that $g_{ij}(q) =
  \delta_{ij}$\footnote{
    Let $E_1, .., E_k$ be an orthornormal basis in $T_qM$. Let $(u_1, .., u_k)$
    be a local coordinates at $q$. Then $E_i = a_{ji}\frac{\partial w}{\partial
      u^j}(q)$, where the matrix $(a_{ij})$ is non singular. Define $(v_1, ..,
    v_k)$ by $u_i = a_{ij}v^j$. Then $\frac{\partial w}{\partial v_k} =
    \frac{\partial w}{\partial u^j}\frac{\partial u^j}{\partial v^i} = E_i$.}
  . The embedding determines analytic coordinate functions
  $w_\alpha(z^1, .., z^k)$, $\alpha = 1, .., n$.  In $\CC^n$, we consider the
  Hermitian inner product $ v \cdot w = \sum_i v_i\bar{w}_i$.

  Let $v$ be a unit vector in $\CC^n$, orthogonal to $M$ at $q$. We consider the
  quadratic forms

  \begin{equation}
    Q(z^1, .., z^k) = 
    \sum_{i,j} v \cdot \frac{\partial^2 w}{\partial z^i \partial z^j}z^iz^j
  \end{equation}
  and

  \begin{equation}
    \label{QQ}
    Q'(x^1, .., x^k, y^1, .., y^k) = Re\{Q(z^1, .., z^k)\}
  \end{equation}
  where $z^j = x^j+iy^j$.

  We have

  \begin{equation}
    \begin{aligned}
      \frac{\partial^2 w}{\partial z^i \partial z^j} &=
      \frac{1}{4}\left(
        \frac{\partial}{\partial x^i}-i\frac{\partial}{\partial
          y^i}\right)\left(\frac{\partial}{\partial
          x^j}-i\frac{\partial}{\partial y^j}
      \right)w \\
      &= \frac{\partial^2 w}{\partial x^i \partial x^j} =
      -\frac{\partial^2 w}{\partial y^i \partial y^j} = 
      -i\frac{\partial^2 w}{\partial x^i \partial y^j} = 
      -i\frac{\partial^2 w}{\partial y^i \partial x^j}
    \end{aligned}
  \end{equation}
  Substituting the above relations into \eqref{QQ}, we find that $Q'$ is exactly
  the second fundamental form of $M$ at $q$ in $\RR^{2n}$.  Recall that the
  focal points are exactly $q+K_i^{-1}v$, where $K_i$ are the principal
  curvatures. From the previous assertion, we see that $-K_i$ is also a
  principal curvature. Thus the focal points lie symmetrically about $q$ on the
  line passing through $q$ in the $v$ direction.
\end{proof}

Now we go back to the proof of the theorem. Choose $p$ in $\CC^n$ such that
$L_p$ is non-degenerate. Since $M$ is a closed subset of $\CC^n$, each $M^a =
L_p^{-1}[0, a]$ is compact. For each $q$ there are at most $2k$ focal points
that lie on line $\stackrel{\leftrightarrow}{pq}$. Since they are distributed
about $q$, there are at most $k$ focal points that lie on the line segment from
$q$ to $p$. By lemma \ref{index}, the index of $L_p$ at $q$ is at most $k$. By
theorem \ref{main}, $M$ has the homotopy type of a CW-complex of dimension at
most $k$. This completes the proof.

\begin{cor}
  If $M \subset \CC^n$ is a non-singular affine algebraic variety in $\CC^n$
  with real dimension $2k$, then
  \begin{equation}
    H_i(M; \ZZ) = 0
  \end{equation}
  for $i>k$.
\end{cor}

\begin{proof}
  $M$ has the homotopy type of a CW-complex of dimension $\le k$. Since $M$ has
  no cells of dimension $ > k$, the $i$th cellular chain complex is $0$.
\end{proof}

\begin{cor}[Lefschetz]
  Let $V$ be an algebraic variety of complex dimension $k$ which lies in the
  complex projective space $\CC P_n$. Let $P$ be a hyperplane in $\CC P_n$ which
  contains the singular points (if any) of $V$. Then the inclusion map
  \begin{equation}
    V\cup P \to V
  \end{equation}
  induces isomorphisms of homology groups in dimensions less than
  $k-1$. Furthermore, the induced homomorphism
  \begin{equation}
    H_k(V\cup P;\ZZ) \to H_{k-1}(V;\ZZ)
  \end{equation}
  is onto.
\end{cor}

\begin{proof}
  The short exact sequence
  \begin{equation}
    \xymatrix{
      0 \ar[r] &
      S_{\cdot}(V\cap P;\ZZ) \ar[r] &
      S_{\cdot}(V;\ZZ) \ar[r] &
      S_{\cdot}(V, V\cap P;\ZZ) \ar[r] & 
      0
    }
  \end{equation}
  induces the long exact sequence
  \begin{equation}
    \xymatrix{
      \cdots \ar[r] &
      H_{r+1}(V, V\cap P;\ZZ) \ar[r] &
      H_{r}(V\cap P;\ZZ) \ar[r] &
      H_{r}(V,;\ZZ) \ar[r] & 
      H_{r}(V, V\cap P;\ZZ) \ar[r] &
      \cdots
    }
  \end{equation}

  Therefore it suffices to show that $H_r(V, V\cap P; \ZZ) = 0$ for $r \le
  k-1$. The Lefschetz duality theorem asserts that
  \begin{equation}
    H_r(V,V\cap P;\ZZ) \cong H^{2k-r}(V-(V\cap P);\ZZ).
  \end{equation}

  Since $V-(V\cap P)$ is a complex analytic manifold embedded as a closed set in
  $\CC^n$, it has the homotopy type of a CW-complex of dimension $\le
  k$. Therefore, by the previous corollary, $H_{2k-r}(V-(V\cap P);\ZZ) = 0$ for
  $r \le k-1$.
\end{proof}

\begin{thebibliography}{9}                                                                                                
%\bibitem{Car} \textsc{Carmo, M. do},\ \textit{Riemannian Geometry}, Birkh\"auser, 1992.

\bibitem{Mil} \textsc{Milnor, J.},\ \textit{Morse Theory}, Princeton University
  Press, 1969.

\bibitem{Hat} \textsc{Hatcher, A.},\ \textit{Algebraic Topology}, Cambridge
  University Press, 2002.

\bibitem{Spa} \textsc{Spanier, E.},\ \textit{Algebraic Topology}, McGraw-Hill,
  1966.

\end{thebibliography}

\end{document}